### 2024-10-28 paper/3-motivation.tex
Update motivation

Example 1 is a WebAssembly code example that has \texttt{loop} and
\texttt{unreachable}
Here, result type of the \texttt{loop} is \texttt{i32} and the \texttt{loop}
has a block of three instructions: \texttt{i32.const}, \texttt{br}, and
\texttt{unreachable}.
\cref{fig:loop} is the \officialp{} of the \texttt{loop} instruction.
It says that the continuation is the start of the loop, the information is
stored in a label, and it \textit{enters} the block with the label.
The term \textit{enter} means that it pushes the label in the stack and makes
pc points to the first instruction in the block.
The \texttt{i32.const} instruction is executed first, which just pushes the
\texttt{i32} value \texttt{42} in the stack.
Then, The \texttt{br} instruction is executed.
\cref{fig:br} is the \officialp{} of the \texttt{br} instruction.
It pops the label from the stack, and makes pc points to the continuation of
the label, which is the start of the loop instruction.
As a result, the code example above is a infinite loop pushing \texttt{42}
forever without executing \texttt{unreachable} instruction.

@@

% exiting label in official prose
There is also a special form of a behavior related to the block: an
\textit{exiting label}.
\cref{fig:exiting-label} is the \officialp of the exiting label.
It is special because this behavior is not performed without an explicit
WebAssembly instruction.
Rather, the behavior is performed when \textbf{the end of a block is reached}
without control instructions or runtime error.
The behavior is that the label is popped from the stack, and the pc changes to
the point after the end of the block.

@@

Example 2 is \texttt{loop} with a block that has only \texttt{i32.const}.
A label is pushed in the stack when a \texttt{loop} instruction is executed.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
Therefore, the code example above results in the value 42 and 3.14 pushed to
the stack.

% general control flow structure of official prose
Similar to the control flow using the block and the label, there are many other
control instructions and structures such as a \texttt{call} instruction that
calls a function with a frame and a \texttt{return} instruction that returns from
the function with the frame.
Not only that, there are the special behaviors such as a \textit{returning from
a function} that corresponds to the exiting label.

% control flow structure in SpecTec prose
However, the \spectecp{} assumes different a model to explain the control flow
of WebAssembly.
Rather than using the notion of pc, \spectecp{} assumes that WebAssembly
instructions are just given one by one.
This is because the \spectecp{} is generated automatically from the
\red{SpecTec dsl}.
\red{SpecTec dsl} uses rewrite rule to describe WebAssembly semantics, and it
can be seen as consuming a WebAssembly instruction and values in the stack,
changing the stack and the store, and inputing new WebAssembly instructions.

@@

Consider example 1 above.
\texttt{loop} and \texttt{unreachable} is the input instruction for the model.
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself and \textit{enters} the block.
Here, \textit{enter} means that it pushes the label in the stack and inputs the
instructions as inputs to execute.
As a result, after the \texttt{loop} is executed, \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, and \texttt{unreachable} are remained as
inputs to the model.
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block, considers the loop instruction in the label as a new input
instruction, and executes it again.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}.
Therefore, it explains the behavior in different point of view, but actually
describes the same thing.

>>>>

Example 1 is a WebAssembly code example that has \texttt{loop} and
\texttt{unreachable}
Here, result type of the \texttt{loop} is \texttt{i32} and the \texttt{loop}
has a block of three instructions: \texttt{i32.const}, \texttt{br}, and
\texttt{unreachable}.
\cref{fig:loop} is the \officialp{} of the \texttt{loop} instruction.
It says that the continuation is the start of the loop, the information is
stored in a label, and it \textit{enters} the block with the label.
The term \textit{enter} means that it pushes the label in the stack and makes
pc points to the first instruction in the block.
The \texttt{i32.const} instruction is executed first, which just pushes the
\texttt{i32} value \texttt{42} in the stack.
Then, The \texttt{br} instruction is executed.
\cref{fig:br} is the \officialp{} of the \texttt{br} instruction.
It pops the label from the stack, and makes pc points to the continuation of
the label, which is the start of the loop instruction.
As a result, the code example above is a infinite loop pushing \texttt{42}
forever without executing \texttt{unreachable} instructions.

@@

% exiting label in official prose
There is also a special form of a behavior related to the block: an
\textit{exiting label}.
\cref{fig:exiting-label} is the \officialp{} of the \textit{exiting label}.
It is special because this behavior is performed without an explicit
WebAssembly instruction.
Rather, the behavior is performed when \textbf{the end of a block is reached}
without control instructions or runtime error.
The behavior is that the label is popped from the stack, and the pc changes to
the point after the end of the block.

@@

Example 2 is a WebAssembly code example that has \texttt{loop} whose block is
only \texttt{i32.const}, and \texttt{f32.const}
When the \texttt{loop} instruction is executed, a label is pushed in the stack,
and it enters the block.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
As a result, after executing the code, the value 42 and 3.14 pushed to the
stack.

% general control flow structure of official prose
Similar to the control flow using the block and the label, there are many other
control instructions and structures such as a \texttt{call} instruction that
calls a function with pushing a frame to stack and a \texttt{return}
instruction that returns from the function with popping the frame from the
stack.
Not only that, there are the special behaviors such as a \textit{returning from
a function} that corresponds to the exiting label.

% control flow structure in SpecTec prose
However, the \spectecp{} assumes different a model to explain the control flow
of WebAssembly.
Rather than using the notion of a pc, \spectecp{} assumes that WebAssembly
instructions are just given one by one.
This is because the \spectecp{} is generated automatically from the
\red{SpecTec dsl}.
\red{SpecTec dsl} uses rewrite rule to describe WebAssembly semantics, and it
can be seen as consuming a WebAssembly instruction and values in the stack,
changing the stack and the store, and inputing new WebAssembly instructions.

@@

Consider example 1 above.

\texttt{loop} and \texttt{unreachable} is the input instruction for the model.
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself and \textit{enters} the block.
Here, \textit{enter} means that it pushes the label in the stack and inputs the
instructions as inputs to execute.
As a result, after the \texttt{loop} is executed, \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, and \texttt{unreachable} are remained as
inputs to the model.
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block, considers the loop instruction in the label as a new input
instruction, and executes it again.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}.
Therefore, it explains the behavior in different point of view, but actually
describes the same thing.
